Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools

### 2024-03-25 paper.qmd
Add content from Declan

<!-- The results for the other -->

@@

CRUSE is a "great tool for getting started with project outline documents". It's an "ideal tool for generating a strategic rationale for estimating demand and building the business case in the absence of cycle count data" (Declan Keenan, Senior Transport Planner working on active travel, personal communication).
"It's a good representation of what's going on"

"I like to know that the data on the county maps is from POWSCAR data which is gold standard".

>>>>

<!-- The results for the other -->
## Selected qualitative results {#sec-qualitative}

Thoughout the project, we undertook a series of workshops and interviews with stakeholders in Ireland to gather feedback on the approach and resulting tool.
The feedback was generally positive, with stakeholders commenting on the need for the tool, lack of data on cycling potential proving a barrier to decision-making and investment, and the potential for the tool.
At an end of project workshop in December 2023, stakeholders were asked to provide feedback on the tool and its potential use in practice.
Quotes and suggestions from these workshops, and a subsequent interview with a Senior Transport Planning using the evidence in their everyday work in Transport Infrastructure Ireland, and presented in this section (with quotes being attributed to named individuals where permission has been granted).

Early in the project, we presented the approach to practitioners in Kildare and Limmerick.
The feedback from the workshop in Kildare highlighted the need for the tool for practitioners at the county level, as highlighted in the following quote:

> "It's a missing piece of evidence that will help new projects get built" (Senior Engineer, Sustainable Transport & Traffic Management, Kildare County Council)

Practitioners in Kildare County Council that the evidence generated *was already valuable* the development of new cycle routes, especially when making the case during planning applications. They emphasized that cycle networks are often fragmented, and the tool allows them to demonstrate how individual local schemes can contribute to a larger, interconnected network. Council officers expressed a preference for visualizations that directly compare the fast and quiet routes, as they recognize the importance of both types of routes (users can switch between the on the 'Route types' page for each county although this functionality is not available in the landing page map due to other feedback requesting a simpler interface).

In comparison with another tool (the National Transport Authority's cycling propensity tool), the CRUSE tool was seen by practitioners in Kildare as more user-friendly and intuitive, with the ability to generate evidence at the route level, which is crucial for planning new cycling infrastructure.
Overall, the workshop participants concluded that the approach filled gap in evidence and will greatly assist in the development of new cycling projects.

During the workshop with practitioners in Limerick, the estimates of Baseline and Go Dutch potential at the network levels were found to match well with local knowledge. The tool effectively highlighted the same sections on the transport network that experienced planners identified as priorities for investment in cycling, such as the South Circular Road. Limerick County Council expressed their interest in utilizing each network layer, particularly the quiet network, to support new cyclists who lack confidence to share space with motor traffic.
Specific suggestions from the workshop included the suggestion to align scenarios with national strategy (resulting in the Climate Action Plan scenario) and to provide data downloads provided as 'Shapefiles' for further analysis (now implemented although the download format is GeoPackage, an open standard for spatial data).

In a subsequent interview with a practitioner using the evidence generated by the CRUSE approach in their everyday work, we received the following feedback.
The outputs were a "great tool for getting started with project outline documents" and "ideal for generating a strategic rationale for estimating demand and building the business case in the absence of cycle count data" (Declan Keenan, Senior Transport Planner working on active travel, personal communication).
Declan also noted the alignment between local knowledge, cycle count data, and the outputs of the approach: "It's a good representation of what's going on" and the networks allign well in terms of network "shape" and the proportion of trips of trips on different parts of the network.
There were two suggestions from Declan which have not yet been implemented: the provision of data showing only "gold standard" POWSCAR data, and the need for training on the tool to ensure that all staff can use it effectively.
These, and other potential improvements, are discussed in the next section.

@@

### 2024-03-25 paper.tex
Add content from Declan

Area-based tools were developed first. Parkin and colleagues presented a
regression model to estimate the percentage of cycling trips to work in
England and Wales based on socioeconomic, transportation, and physical
factors at the small area (wards) level in 2007
\citep{parkin_estimation_2007}. They found that areas with a higher
percentage of females, non-whites, car ownership, lower socioeconomic
classes, income, distance to work, population density, poor highway
conditions, hills, rainfall, and fewer off-road bicycle routes tended to
have a lower proportion of cycling to work.

@@

The main result of the work presented in this paper is an open access
web application and datasets on current and potential future cycling
levels in Ireland, with evidence provided at county and route segment
levels. Like the PCT on which CRUSE builds, a key feature of the results
is that they are systematic and available for all counties, and most
routes on which cycling is permitted, in Ireland. The full set of
results are therefore too extensive to present in their entirety in this
paper. Through the interactive web application, users can explore the
data to generate the results that are most relevant to their needs,
whether that is finding the cycling potential on a particular road or
finding `weak links' or barriers in the cycle network associated with a
particular school, work place or other destination.

Instead of trying to show such use cases, of which there are many
hundreds, we present a selection of results to illustrate the main
features of the CRUSE Tool.

@@

@@

CRUSE is a ``great tool for getting started with project outline
documents''. It's an ``ideal tool for generating a strategic rationale
for estimating demand and building the business case in the absence of
cycle count data'' (Declan Keenan, Senior Transport Planner working on
active travel, personal communication). ``It's a good representation of
what's going on''

``I like to know that the data on the county maps is from POWSCAR data
which is gold standard''.

>>>>

Area-based approaches were developed before network-based tools, such as
a 2007 regression-based model to estimate the percentage of cycling
trips to work in England and Wales based on socioeconomic,
transportation, and physical factors at the small area (wards) level in
2007 \citep{parkin_estimation_2007}. They found that areas with a higher
percentage of females, non-whites, car ownership, lower socioeconomic
classes, income, distance to work, population density, poor highway
conditions, hills, rainfall, and fewer off-road bicycle routes tended to
have a lower proportion of cycling to work. The approach also generated
``forecasts for potential levels of bicycle use'' with with increases in
off-road routes, improvements of highway and, conversely, increases in
car ownership and distance to work leading to decreases in cycling to
work.

@@

The main result of the work presented in this paper is an open access
web application and datasets on current and potential future cycling
levels in Ireland, with evidence provided at county and route segment
levels. Like other national tools presented in Table~\ref{tbl-tools}, a
key feature of the results is that it provides a consistent and
objective baseline for comparing cycling potentials in different areas.

The approach generates estimates of cycling potential on every road in
the country and are therefore too extensive to present in their entirety
in this paper. Through the interactive web application, users can
explore the data to generate the results that are most relevant to their
needs, whether that is finding the cycling potential on a particular
road or finding `weak links' or barriers in the cycle network associated
with a particular school, work place or other destination.

Instead of trying to show such use cases, of which there are many
hundreds, we present a selection of results to illustrate the main
features of the resulting evidence, in Section~\ref{sec-quantiative}. We
also present some qualitative results based on workshops and interviews
with stakeholders in Ireland, in Section~\ref{sec-qualitative}.

\subsection{Selected quantitative results}\label{sec-quantiative}

@@

\subsection{Selected qualitative results}\label{sec-qualitative}

Thoughout the project, we undertook a series of workshops and interviews
with stakeholders in Ireland to gather feedback on the approach and
resulting tool. The feedback was generally positive, with stakeholders
commenting on the need for the tool, lack of data on cycling potential
proving a barrier to decision-making and investment, and the potential
for the tool. At an end of project workshop in December 2023,
stakeholders were asked to provide feedback on the tool and its
potential use in practice. Quotes and suggestions from these workshops,
and a subsequent interview with a Senior Transport Planning using the
evidence in their everyday work in Transport Infrastructure Ireland, and
presented in this section (with quotes being attributed to named
individuals where permission has been granted).

Early in the project, we presented the approach to practitioners in
Kildare and Limmerick. The feedback from the workshop in Kildare
highlighted the need for the tool for practitioners at the county level,
as highlighted in the following quote:

\begin{quote}
``It's a missing piece of evidence that will help new projects get
built'' (Senior Engineer, Sustainable Transport \& Traffic Management,
Kildare County Council)
\end{quote}

Practitioners in Kildare County Council that the evidence generated
\emph{was already valuable} the development of new cycle routes,
especially when making the case during planning applications. They
emphasized that cycle networks are often fragmented, and the tool allows
them to demonstrate how individual local schemes can contribute to a
larger, interconnected network. Council officers expressed a preference
for visualizations that directly compare the fast and quiet routes, as
they recognize the importance of both types of routes (users can switch
between the on the `Route types' page for each county although this
functionality is not available in the landing page map due to other
feedback requesting a simpler interface).

In comparison with another tool (the National Transport Authority's
cycling propensity tool), the CRUSE tool was seen by practitioners in
Kildare as more user-friendly and intuitive, with the ability to
generate evidence at the route level, which is crucial for planning new
cycling infrastructure. Overall, the workshop participants concluded
that the approach filled gap in evidence and will greatly assist in the
development of new cycling projects.

During the workshop with practitioners in Limerick, the estimates of
Baseline and Go Dutch potential at the network levels were found to
match well with local knowledge. The tool effectively highlighted the
same sections on the transport network that experienced planners
identified as priorities for investment in cycling, such as the South
Circular Road. Limerick County Council expressed their interest in
utilizing each network layer, particularly the quiet network, to support
new cyclists who lack confidence to share space with motor traffic.
Specific suggestions from the workshop included the suggestion to align
scenarios with national strategy (resulting in the Climate Action Plan
scenario) and to provide data downloads provided as `Shapefiles' for
further analysis (now implemented although the download format is
GeoPackage, an open standard for spatial data).

In a subsequent interview with a practitioner using the evidence
generated by the CRUSE approach in their everyday work, we received the
following feedback. The outputs were a ``great tool for getting started
with project outline documents'' and ``ideal for generating a strategic
rationale for estimating demand and building the business case in the
absence of cycle count data'' (Declan Keenan, Senior Transport Planner
working on active travel, personal communication). Declan also noted the
alignment between local knowledge, cycle count data, and the outputs of
the approach: ``It's a good representation of what's going on'' and the
networks allign well in terms of network ``shape'' and the proportion of
trips of trips on different parts of the network. There were two
suggestions from Declan which have not yet been implemented: the
provision of data showing only ``gold standard'' POWSCAR data, and the
need for training on the tool to ensure that all staff can use it
effectively. These, and other potential improvements, are discussed in
the next section.

@@
